---
title: "Practical Machine Learning"
author: "Artur Mrozowski"
date: "28. jan. 2016"
output: 
  html_document: 
    keep_md: yes
---

### Introduction  
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.  

In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.  


```{r, echo=FALSE}
library(caret)
library(ggplot2)
library(rpart)
library(rpart.plot)
```

## Read the data

```{r, echo=TRUE}
trainUrl<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train<-read.csv(trainUrl)
test<-read.csv(testUrl)
```
##Clean the data
The columns that contain NA values need to be removed. Also the columms that don't make much sensne in the prediction model and create unecessary noise.
```{r, echo=TRUE}
#Get rid of not complete cases
train <- train[, colSums(is.na(train)) == 0] 
test <- test[, colSums(is.na(test)) == 0] 
#Get Rid of not importan variables
classe <- train$classe
trainRmv <- grepl("^X|timestamp|window", names(train))
train <- train[, !trainRmv]
trainCln <- train[, sapply(train, is.numeric)]
trainCln$classe <- classe
testRmv <- grepl("^X|timestamp|window", names(test))
test <- test[, !testRmv]
testCln <- test[, sapply(test, is.numeric)]

```
##Create training and test data sets
By splitting the data into training(70%) and validation(30%) we'll be able to validate the model.  
```{r, echo=TRUE}
set.seed(3334)
inTrain <- createDataPartition(trainCln$classe, p=0.70, list=F)
trainDataSet <- trainCln[inTrain, ]
testDataSet <- trainCln[-inTrain, ]


```
#Train the model randomforest
In random forests, there is no need for cross-validation . It is estimated internally, during the run, as follows:

Each tree is constructed using a different bootstrap sample from the original data. About one-third of the cases are left out of the bootstrap sample and not used in the construction of the kth tree.   
```{r, echo=TRUE}
modelRf <- train(classe ~ ., data=trainDataSet, method="rf", ntree=250)
modelRf
```
#Variable importance
```{r}
varImp(modelRf)
ggplot(varImp(modelRf))

```

we can see that rooll_belt and yaw_belt  are by far the most important variables

#accuracy plot
```{r}
plot.train(modelRf)
```

Accuracy is 99,4%

Let's evaluate model results in confusion matrix after estimating the performance on the validation data set
```{r, echo=TRUE}
predRf<-predict(modelRf,testDataSet)
confusionMatrix(testDataSet$classe,predRf)
```

Let's reduce number of variable to the most importan ones. Let's reduce training and testing set accordingly and validation set on which we make the predictions for the classe variable.
```{r}

trainReduced<-trainDataSet[,c("roll_belt","yaw_belt","magnet_dumbbell_z","pitch_belt","magnet_dumbbell_y","magnet_dumbbell_x","pitch_forearm","classe")]
testReduced<-testDataSet[,c("roll_belt","yaw_belt","magnet_dumbbell_z","pitch_belt","magnet_dumbbell_y","magnet_dumbbell_x","pitch_forearm","classe")]
testClnReduced<-testCln[,c("roll_belt","yaw_belt","magnet_dumbbell_z","pitch_belt","magnet_dumbbell_y","magnet_dumbbell_x","pitch_forearm")]

```

Now let's train the model on the reduced data set.
```{r}
modelRfReduced <- train(classe ~ ., data=trainReduced, method="rf", ntree=250)
```
Prediciton for the new model with reduced number of variables.
```{r}
predReduced<-predict(modelRfReduced,testReduced)
confusionMatrix(testReduced$classe,predReduced)
oose <- 1 - as.numeric(confusionMatrix(testReduced$classe,predReduced)$overall[1])
oose
```

Out of sample error
```{r}
oose <- 1 - as.numeric(confusionMatrix(testReduced$classe,predReduced)$overall[1])
oose
```


#Results

```{r}
resReduced<-predict(modelRfReduced,testClnReduced)
resReduced
```

#Visualizaiton of the tree model
```{r}
treeModel <- rpart(classe ~ ., data=trainReduced, method="class")
 prp(treeModel)
 
```

